---
title: "Elden Ring Weapon Analysis"
author: "Nathan Villicaña-Shaw"
date: "`r Sys.Date()`"
output: html_document
---

## Inital environment setup
```{r setup}
# Set CRAN mirror
options(repos = c(CRAN = "http://cran.utstat.utoronto.ca/"))
# we will need the dplyr package for mutate() and data frame manipulation
install.packages("dplyr")
install.packages("jsonlite")
# The dplyr package provides a grammar of data manipulation, 
# providing a consistent set of verbs that help you solve the 
# most common data manipulation challenges: 
# filtering rows, selecting columns, rearranging rows and 
# columns, summarizing data, and creating new variables.
library("dplyr")

# The ggplot2 package is a system for declaratively creating 
# graphics, based on The Grammar of Graphics. 
# It provides a range of tools for constructing plots, 
# from basic plotting of data to highly sophisticated and 
# customized visualizations.
library("ggplot2")

# The purrr package is a functional programming toolkit 
# for R. It provides a complete and consistent set of tools 
# for working with functions and vectors. It's main purpose 
# is to allow you to work with data in a way that is more 
# consistent with functional programming principles.
library("purrr")

# The tidyr package provides a set of functions to help 
# organize messy data into a structured and tidy form. 
# Tidy data has a specific structure where each variable 
# is a column, each observation is a row, and each type of 
# observational unit is a separate table.
library("tidyr")

# The jsonlite package provides functions to convert R objects 
# to and from JSON format, which is a lightweight data-interchange 
# format that is easy for humans to read and write, and easy for 
# machines to parse and generate. It is commonly used for 
# transmitting data between a server and a web application, 
# as an alternative to XML.
library("jsonlite")

```

## Pull in weapon data from .CSV file
```{r}
setwd("../data/elden_ring")
# if this file is not loading change its permissions using chmod =)
all_weapons <- read.csv("weapons.csv")
colnames(all_weapons)
```
## Initial investigation of the CSV file
First lets probe the data a little bit to understand how it is being stored...
```{r all_weapons_summary}
summary(all_weapons)
```
## Data Cleaning
Okay, cool, now lets rename our columns so they make more sense
```{r}
# rename columns to something more human friendly
all_weapons <- all_weapons %>% 
  rename(Category = Type,
         Physical_Damage = Phy,
         Magic_Damage = Mag,
         Fire_Damage = Fir,
         Lightning_Damage = Lit,
         Holy_Damage = Hol,
         Critical_Scaler = Cri,
         Stamina_Drain = Sta,
         Strength_Scaler = Str,
         Dexterity_Scaler = Dex,
         Intelligence_Scaler = Int,
         Faith_Scaler = Fai,
         Arcane_Scaler = Arc,
         Special_Effect_Damage = Any,
         Physical_Block_Damage = Phy.1,
         Magic_Block_Damage = Mag.1,
         Fire_Block_Damage = Fir.1,
         Lightning_Block_Damage = Lit.1,
         Holy_Block_Damage = Hol.1,
         Boost = Bst,
         Rest = Rst,
         Weight = Wgt,
         Upgrade_Items = Upgrade) %>%
  mutate(across(Physical_Damage:Stamina_Drain, ~ifelse(. == "-", "0", .))) %>%
  mutate(across(Special_Effect_Damage:Rest, ~ifelse(. == "-", "0", .)))

# sweet, lets check out our newly named data frame
```


Next we need to replace any NA and Nulls in the table with 0's
```{r}
str(all_weapons)
```
We also have to convert many of the damage numbers into ints from strings and create a column for Total_Damage and Total_Block_Damage

```{r}
# This code uses the mutate() function along with across() to apply the as.integer() function to multiple columns of all_weapons at once. The resulting dataframe will have the specified columns converted from strings to integers.
all_weapons <- all_weapons %>%
  mutate(across(c(Physical_Damage, Magic_Damage, Fire_Damage, Lightning_Damage, Holy_Damage, Stamina_Drain, Special_Effect_Damage, Physical_Block_Damage, Magic_Block_Damage, Fire_Block_Damage, Lightning_Block_Damage, Holy_Block_Damage, Boost, Rest), as.integer))

## Create a new Column for "total damage" as a sum of all damage types
all_weapons <- all_weapons %>% 
  mutate(Total_Damage = Physical_Damage + Magic_Damage + Fire_Damage + Lightning_Damage + Holy_Damage) %>%
  relocate(Total_Damage, .after = "Category") %>% 
  mutate(Total_Block_Damage = Physical_Block_Damage + Magic_Block_Damage + Fire_Block_Damage + Lightning_Block_Damage + Holy_Block_Damage) %>%
  relocate(Total_Block_Damage, .after = "Special_Effect_Damage")

head(all_weapons)
```

We might want to consider replacing the somber and regular smithing stones with 0 and 1 later on, but this should be good for now

# Some initial plots
```{r}
ggplot(all_weapons, aes(x=Weight, y=Physical_Damage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~Category) + 
  labs(title="Comparison of Total Damage and Weight for Elden Ring Weapons")
```
## Investigating purely physical weapons
```{r}
# filter the weaponset to only include weapons with physical damage
# this is true if total damage equals physial damage
physical_weapons <- all_weapons %>% filter(., Physical_Damage == Total_Damage, Weight > 0, Physical_Damage > 0)
print(length(physical_weapons))

ggplot(physical_weapons, aes(x=Weight, y=Physical_Damage)) +
  geom_point() + 
  geom_smooth(method = "lm", formula = y ~ log(x), se=FALSE) +
  scale_y_log10() +
  guides(color=FALSE) +
  labs(title="Comparison of Physical Damage and Weight for Elden Ring Weapons")
```

## Convert Damage Scaling to Integers
To make our analysis easier, lets map our letter values for the Damage Scaling values to numbers
- = 0
E = 1
D = 2
C = 3
B = 4
A = 5
S = 6

Also, lets map our smithing_stones to 0 and somber_smithing_stones to 1

```{r}
# remap our values as strings
all_weapons <- all_weapons %>% 
  mutate(across(Strength_Scaler:Arcane_Scaler, ~ifelse(. == "-", "0", .))) %>%
  mutate(across(Strength_Scaler:Arcane_Scaler, ~ifelse(. == "E", "1", .))) %>%
  mutate(across(Strength_Scaler:Arcane_Scaler, ~ifelse(. == "D", "2", .))) %>%
  mutate(across(Strength_Scaler:Arcane_Scaler, ~ifelse(. == "C", "3", .))) %>%
  mutate(across(Strength_Scaler:Arcane_Scaler, ~ifelse(. == "B", "4", .))) %>%
  mutate(across(Strength_Scaler:Arcane_Scaler, ~ifelse(. == "A", "5", .))) %>%
  mutate(across(Strength_Scaler:Arcane_Scaler, ~ifelse(. == "S", "6", .))) %>%
  mutate(across(Upgrade_Items, ~ifelse(. == "Somber Smithing Stones", "1", .))) %>%
  mutate(across(Upgrade_Items, ~ifelse(. == "Smithing Stones", "0", .)))

# convert our string values to ints
all_weapons <- all_weapons %>%
  mutate(across(c(Strength_Scaler, Dexterity_Scaler, Faith_Scaler, Intelligence_Scaler, Arcane_Scaler), as.integer))

head(all_weapons)
```

# Plotting Relationship between Total_Blocking_Damage and Total_Damage

```{r}
weapons_with_block_damage <- filter(all_weapons, Total_Block_Damage > 0)

ggplot(weapons_with_block_damage, aes(x=Total_Damage, y=Total_Block_Damage, color=Weight, shape=Upgrade_Items)) +
  geom_point() + 
  geom_smooth(method = "lm", se=FALSE) +
  scale_y_log10() +
  guides(color=FALSE) +
  facet_wrap(~Upgrade_Items) +
  labs(title="Comparison of Total Damage (x), Total Block Damage (y), and Weight (color)")
```

## Splitting Apart Staffs, and Seals into a group, Bows and Crossbows into a group, and all other weapons in final group

```{r}
magic_catalists <- filter(all_weapons, Category == "Glintstone Staff" | Category == "Sacred Seal")
ranged_weapons <- filter(all_weapons, Category == "Bow" | Category == "Ballista" | Category == "Crossbow" | Category == "Greatbow" | Category == "Light Bow")
melee_weapons <- filter(all_weapons, Category != "Glintstone Staff" | Category != "Sacred Seal" | Category != "Bow" | Category != "Ballista" | Category != "Crossbow" | Category != "Greatbow" | Category != "Light Bow")

head(magic_catalists)
head(ranged_weapons)
head(melee_weapons)
summary(magic_catalists)
summary(ranged_weapons)
summary(melee_weapons)
```

### Plotting the different categories separtely

```{r}
ggplot(melee_weapons, aes(x=Weight, y=Total_Damage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Total Damage (y) and Weight (x) for Elden Ring Melee Weapons")

ggplot(ranged_weapons, aes(x=Weight, y=Total_Damage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Total Damage (y) and Weight (x) for Elden Ring Ranged Weapons")

ggplot(magic_catalists, aes(x=Weight, y=Total_Damage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Comparison of Total Damage and Weight for Elden Ring Magic Catalists")
```

### Taking a look at comparing the attack damage and block damage for the weapons
```{r}
ggplot(melee_weapons, aes(x=Total_Block_Damage, y=Total_Damage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Total Damage (y) and Weight (x) for Elden Ring Melee Weapons")

ggplot(ranged_weapons, aes(x=Total_Block_Damage, y=Total_Damage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Total Damage (y) and Weight (x) for Elden Ring Ranged Weapons")

ggplot(magic_catalists, aes(x=Total_Block_Damage, y=Total_Damage)) +
  geom_point() + 
  geom_smooth(method = "lm", se = FALSE) +
  labs(title="Comparison of Total Damage and Weight for Elden Ring Magic Catalists")
```
## Sweet, now lets check out Dex scaled weapons
```{r}
dex_scaled_weapons <- filter(all_weapons, Dexterity_Scaler > 0)
str_scaled_weapons <- filter(all_weapons, Strength_Scaler > 0)
arc_scaled_weapons <- filter(all_weapons, Arcane_Scaler > 0)
int_scaled_weapons <- filter(all_weapons, Intelligence_Scaler > 0)
fai_scaled_weapons <- filter(all_weapons, Faith_Scaler > 0)

head(dex_scaled_weapons)
head(str_scaled_weapons)
head(arc_scaled_weapons)
head(int_scaled_weapons)
head(fai_scaled_weapons)
length(arc_scaled_weapons)
length(fai_scaled_weapons)
```

```{r}
df <- data.frame(
  Scale_Type = c("Dexterity", "Strength", "Arcane", "Intelligence", "Faith"),
  Count = c(length(dex_scaled_weapons),
            length(str_scaled_weapons),
            length(arc_scaled_weapons),
            length(int_scaled_weapons),
            length(fai_scaled_weapons))
)

df <- df[order(df$Scale_Type),]  # Sort the data frame by Scale_Type

ggplot(df, aes(x = "", y = Count, fill = Scale_Type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Count, y = cumsum(Count) - (0.5 * Count)), color = "white") +
  coord_flip() +
  labs(title = "Weapons by Scale Type", x = NULL, y = NULL) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

```
## Very Interesting, now lets see what the total scale values for all these categories...
```{r}
df <- data.frame(
  Scale_Type = c("Dexterity", "Strength", "Arcane", "Intelligence", "Faith"),
  Count = c(length(dex_scaled_weapons),
            length(str_scaled_weapons),
            length(arc_scaled_weapons),
            length(int_scaled_weapons),
            length(fai_scaled_weapons)),
  
  Scale_Totals = c(sum(dex_scaled_weapons$Dexterity_Scaler),
                   sum(str_scaled_weapons$Strength_Scaler),
                   sum(arc_scaled_weapons$Arcane_Scaler),
                   sum(int_scaled_weapons$Intelligence_Scaler),
                   sum(fai_scaled_weapons$Faith_Scaler)
            )
  )

ggplot(df, aes(x = Scale_Type, y = Scale_Totals, fill = Scale_Type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Scale_Totals, y = cumsum(Count) - (0.5 * Count)), color = "white") +
  coord_flip() +
  labs(title = "Total Scale Weight", x = NULL, y = NULL) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

ggplot(df, aes(x = Scale_Type, y = Scale_Totals / Count, fill = Scale_Type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Scale_Totals / Count, y = cumsum(Count) - (0.5 * Count)), color = "white") +
  coord_flip() +
  labs(title = "Average Scale Weight", x = NULL, y = NULL) +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()
```

# Hmm, there is something odd going on with the Arcane Weapons....

```{r}
library(dplyr)

arc_weap_summary <- arc_scaled_weapons %>%
  count(Arcane_Scaler) %>%
  mutate(Scale_Amount = factor(Arcane_Scaler,
                               levels = c("1", "2", "3", "4", "5", "6"),
                               labels = c("E", "D", "C", "B", "A", "S"))) %>%
  select(Scale_Amount, n)

ggplot(arc_weap_summary, aes(x = Scale_Amount, y = n)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = n), vjust = -0.5) +
  labs(title = "Arcane Weapons by Scale Amount") +
  theme_minimal()

```